Modeling Ionic Polymer-Metal Composites\\ with Space-Time Adaptive Multimesh \emph{hp}-FEM

### 2014-07-01 results.tex

An example of the solution at $t=0.1\ s$ and $t=3.0\ s$ 
calculated with the HP\_ANISO refinement mode is shown
in Figs.~\ref{fig:cphi-1}~and~\ref{fig:cphi-2}.

@@

\subsection{Comparison of single-mesh and multi-mesh $hp$-FEM}
% David: Error convergence plots here.
Running the simulation with different adaptivity modes 
and meshes showed that the multi-mesh $hp$-FEM configuration resulted in
the smallest problems and similar error 
convergence compared to any single-mesh configuration. However,
multi-mesh problems generally resulted in longer computing times. This is a known
shortcoming of Hermes at this point and it is due to the fact that
multi-mesh uses the union mesh (see Section~\ref{sec:hermes}) where 
the numerical integration of high order is done on very small elements.
The problem size and computing time are illustrated for HP\_ANISO adaptivity mode in Fig.~\ref{fig:singlemultidof} and Fig.~\ref{fig:singlemulticpu}.
The same holds true for HP\_ISO mode.  It must be also noted that the error
converged to or below $0.5\%$ for all $p$-FEM and anisotropic $hp$-FEM results.

@@

\noindent
Figs.~\ref{fig:isoanisopdof} and \ref{fig:isoanisopcpu} present a similar 
comparsion for the P\_ISO, P\_ANISO, and HP\_ANISO\_P modes. Recall that these 
computations use a different initial mesh that was a-priori refined in space.

@@

In real physics calculations, the applied voltage on boundary $\partial\Omega_1$
is not constant. This can be, for instance, due to the high resistance of
the electrodes as explained in~\cite{pugal2009}.
To see how the HP\_ANISO adaptivity works for such situations, the
voltage on the boundary was applied as follows:
\begin{equation}
  \phi_{\Omega_1}\left( x \right)=0.5\left[V \right] \frac{x\left[ m \right ]}{\text{width}_{\Omega_1}\left[ m \right]}+0.5\left[ V \right],
\end{equation}
where $\text{width}_{\Omega_1}$ is the width of the boundary. The given boundary is effectively
a linear incrase of the voltage from $\phi_{\Omega_1}\left(x = 0 \right)=0.5\ V$ to
$\phi_{\Omega_1}\left(x=\text{width}_{\Omega_1}\right) = 1.0\ V$.
Now the concentration gradient $\nabla c$ and the voltage gradient $\nabla \varphi$ are no
longer effectively in 1D.

The calculated scaled values $c$ and $\varphi$ in $\Omega$ and corresponding meshes and polynomial
degrees of the elements at $t=0.1\ s$ are shown in Fig.~\ref{fig:cphi-orders}.
Notice that the solution
is different to the one in Fig.~\ref{fig:cphi-1}. The HP\_ANISO
adaptivity algorithm has particularly increased the polynomial degree
and refined the mesh near $\Omega_1$ where a sharp concentration
peak exists (compare to Fig.~\ref{fig:poly}).
At $t=3.0\ s$, the shape of the solutions $c$ and $\varphi$ are similar to the one
in Fig.~\ref{fig:cphi-2} and therefore the polynomial space and mesh gets adapted
accordingly. This example clearly illustrates how 
the solution of PNP with non-uniform
boundary conditions is very dynamic in time
and how the HP\_ANISO time dependent adaptivity 
finds an optimal mesh and polynomial space to adapt to the dynamics
of the problem.

@@

\begin{figure}[!ht]
  \begin{centering}
  \includegraphics[width=\columnwidth]{debye_cpu}
  \caption{\label{fig:debye_cpu} CPU time depending on the
	Debye length. Initially the coarse mesh (shown in Fig.~\ref{fig:mesh}~(a))
	was used. It was necessary to use the fine mesh (see Fig.~\ref{fig:mesh}~(c)) 
	for smaller $\lambda_D$ values.}
  \end{centering}
\end{figure}
The Debye length $\lambda_D$ is the screening length in the electrolyte
solutions. Its numerical value shows the
thickness of the charged layer in the vicinity of the boundaries $\partial\Omega_1$ 
and $\partial\Omega_2$.
In all the previous simulations, the Debye screening length was determined by  the constants in
Table~\ref{Table:used-constants} and Eq.~\ref{eq:debye}: $\lambda_D=1.7\ \mu m$. It is known that computation gets increasingly
difficult when reducing the value of $\lambda_D$.
It was our interest to see how small screening lengths can Hermes HP\_ANISO automatic
adaptivity handle. The parameter $\varepsilon$ was varied as follows:
$$\varepsilon_n=\varepsilon\times 0.5^n,$$ where $\varepsilon$ is taken from Table~\ref{Table:used-constants}. The simulations were run for each $\varepsilon_n$ and corresponding $\lambda_D$ value
and maximum number of degrees of freedom and cumulative CPU time were recorded.
The simulation time $t$ for each $\lambda_D$ was chosen to be $\tau$ --- the
characteristic time scale --- and each simulation
was divided equally into fifteen time steps. The PID controller was not used.
Fig.~\ref{fig:debye_dof} shows the maximum and average number of degrees of 
freedom during calculation
as a function of the Debye length and Fig.~\ref{fig:debye_cpu} shows cumulative CPU time as a function
of the Debye length. The simulations up to $0.52\ nm$
screening length were carried out on
the initial coarse mesh. However, from $\lambda_D > 0.52\ nm$, the finer initial mesh had to
be used so the existence of the large gradients of the
physical fields $c$ and $\varphi$ near the boundaries could be captured in the first place.
\begin{figure}[!ht]
  \begin{centering}
  \includegraphics[width=\columnwidth]{cphitau}
  \caption{\label{fig:cphitau} Calculated fields $c$ and $\varphi$ at $t=\tau=0.81\ ms$
	for $\lambda_D=0.4\ nm$.}
  \end{centering}
\end{figure}
\noindent
The fine mesh allowed simulations with the Debye length down to $0.40\ nm$~m. The calculated
$c$ and $\varphi$ at $t=\tau$ for $\lambda_D=0.40\ nm$ are shown in Fig.~\ref{fig:cphitau}.
It appears that when using even finer initial mesh and higher initial polynomial degrees, even
smaller Debye lengths could be used when necessary. The polynomial space of $c$ had
consistently higher maximum polynomial degree than that of $\varphi$, however, the difference
was less noticeable for smaller Debye lengths.

>>>>

An example of the solution at $t=0.1$~s and $t=3.0$~s
calculated with the HP\_ANISO refinement mode is shown
in Figs.~\ref{fig:cphi-1}~and~\ref{fig:cphi-2}.

@@

\subsection{Comparison of single-mesh and multi-mesh $hp$-FEM}
Running the simulation with different adaptivity modes 
and meshes showed that the multi-mesh $hp$-FEM configuration resulted in
the smallest problems and similar error 
convergence compared to any single-mesh configuration. However,
multi-mesh problems generally resulted in longer computing times. This is a known
shortcoming of Hermes at this point and it is due to the fact that
multi-mesh uses the union mesh (see Section~\ref{sec:hermes}) where 
the numerical integration of high order is done on very small elements.
The problem size and computing time are illustrated for HP\_ANISO adaptivity mode in Fig.~\ref{fig:singlemultidof} and Fig.~\ref{fig:singlemulticpu}.
The same holds true for HP\_ISO mode.  It must be also noted that the error
converged to or below $0.5\%$ for all $p$-FEM and anisotropic $hp$-FEM results.

@@

\noindent
Figs.~\ref{fig:isoanisopdof} and \ref{fig:isoanisopcpu} present a similar 
comparison for the P\_ISO, P\_ANISO, and HP\_ANISO\_P modes. Recall that these 
computations use a different initial mesh that was a-priori refined in space.

@@

In real physics calculations, the applied voltage on boundary $\partial\Omega_1$
is not constant. This can be, for instance, due to the high resistance of
the electrodes as explained in~\cite{pugal2009}.
To see how the HP\_ANISO adaptivity works for such situations, the
voltage on the boundary was applied as follows:
\begin{equation}
  \phi_{\Omega_1}\left( x \right)=0.5\left[V \right] \frac{x\left[ m \right ]}{\text{width}_{\Omega_1}\left[ m \right]}+0.5\left[ V \right],
\end{equation}
where $\text{width}_{\Omega_1}$ is the width of the boundary. The given boundary is effectively
a linear increase of the voltage from $\phi_{\Omega_1}\left(x = 0 \right)=0.5$~V to
$\phi_{\Omega_1}\left(x=\text{width}_{\Omega_1}\right) = 1.0$~V.
Now the concentration gradient $\nabla c$ and the voltage gradient $\nabla \varphi$ are no
longer effectively in 1D.

The calculated scaled values $c$ and $\varphi$ in $\Omega$ and corresponding meshes and polynomial
degrees of the elements at $t=0.1$~s are shown in Fig.~\ref{fig:cphi-orders}.
Notice that the solution
is different to the one in Fig.~\ref{fig:cphi-1}. The HP\_ANISO
adaptivity algorithm has particularly increased the polynomial degree
and refined the mesh near $\Omega_1$ where a sharp concentration
peak exists (compare to Fig.~\ref{fig:poly}).
At $t=3.0$~s, the shape of the solutions $c$ and $\varphi$ are similar to the one
in Fig.~\ref{fig:cphi-2} and therefore the polynomial space and mesh gets adapted
accordingly. This example clearly illustrates how 
the solution of PNP with non-uniform
boundary conditions is very dynamic in time
and how the HP\_ANISO time dependent adaptivity 
finds an optimal mesh and polynomial space to adapt to the dynamics
of the problem.

@@

\begin{figure}[!ht]
  \begin{centering}
  \includegraphics[width=\columnwidth]{debye_cpu}
  \caption{\label{fig:debye_cpu} CPU time depending on the
	Debye length. Initially the coarse mesh (shown in Fig.~\ref{fig:mesh}~(a))
	was used. It was necessary to use the fine mesh (see Fig.~\ref{fig:mesh}~(c)) 
	for smaller $\lambda_D$ values.}
  \end{centering}
\end{figure}
The Debye length $\lambda_D$ is the screening length in the electrolyte
solutions. Its numerical value shows the
thickness of the charged layer in the vicinity of the boundaries $\partial\Omega_1$ 
and $\partial\Omega_2$.
In all the previous simulations, the Debye screening length was determined by  the constants in
Table~\ref{Table:used-constants} and Eq.~\eqref{eq:debye}: $\lambda_D=1.7\ \mu m$. It is known that computation gets increasingly
difficult when reducing the value of $\lambda_D$.
It was our interest to see how small screening lengths can Hermes HP\_ANISO automatic
adaptivity handle. The parameter $\varepsilon$ was varied as follows:
$$\varepsilon_n=\varepsilon\times 0.5^n,$$ where $\varepsilon$ is taken from Table~\ref{Table:used-constants}. The simulations were run for each $\varepsilon_n$ and corresponding $\lambda_D$ value
and maximum number of degrees of freedom and cumulative CPU time were recorded.
The simulation time $t$ for each $\lambda_D$ was chosen to be $\tau$ --- the
characteristic time scale --- and each simulation
was divided equally into fifteen time steps. The PID controller was not used.
Fig.~\ref{fig:debye_dof} shows the maximum and average number of degrees of 
freedom during calculation
as a function of the Debye length and Fig.~\ref{fig:debye_cpu} shows cumulative CPU time as a function
of the Debye length. The simulations up to $0.52$~nm
screening length were carried out on
the initial coarse mesh. However, from $\lambda_D > 0.52$~nm, the finer initial mesh had to
be used so the existence of the large gradients of the
physical fields $c$ and $\varphi$ near the boundaries could be captured in the first place.
\begin{figure}[!ht]
  \begin{centering}
  \includegraphics[width=\columnwidth]{cphitau}
  \caption{\label{fig:cphitau} Calculated fields $c$ and $\varphi$ at $t=\tau=0.81\ ms$
	for $\lambda_D=0.4\ nm$.}
  \end{centering}
\end{figure}
\noindent
The fine mesh allowed simulations with the Debye length down to $0.40$~nm. The calculated
$c$ and $\varphi$ at $t=\tau$ for $\lambda_D=0.40$~nm are shown in Fig.~\ref{fig:cphitau}.
It appears that when using even finer initial mesh and higher initial polynomial degrees, even
smaller Debye lengths could be used when necessary. The polynomial space of $c$ had
consistently higher maximum polynomial degree than that of $\varphi$, however, the difference
was less noticeable for smaller Debye lengths.
